Visualization and analysis of probability distributions of large temporal data

### 2022-02-09 Rmd/04-gracsr.Rmd
Ch 4 Q1 Juergen
Ch 4 all done Juergen

There is an extensive body of literature focused on time series clustering related to smart meter data. @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics. <!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means [@Rhodes2014-it]. $k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data. To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering [@Ndiaye2011-pf]. PCA eliminates correlation patterns and decreases feature space, but loses interpretability. Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids. Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb. Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency. @Motlagh2019-yj suggested that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered. @chicco2010renyi addressed information theory-based clustering such as Shannon or Renyi entropy and its variations. @Melnykov2013-sp discussed how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions. None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.

@@

```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for our method for clustering time series.", out.width="100%"}
knitr::include_graphics("img/flowchart.png")

@@

(ref:hod-ind-group-png) The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots. They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer. The number indicates a unique customer id and a prototype id.

>>>>

There is an extensive body of literature focused on time series clustering related to smart meter data. @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics. <!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means [@Rhodes2014-it]. $k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data. To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering [@Ndiaye2011-pf]. PCA eliminates correlation patterns and decreases feature space, but loses interpretability. Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids. Many techniques mentioned in @Tureczek2017-pb fail to recognize smart meter readings as a data type with a temporal component [@Tureczek2018-ha]. Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency. @Motlagh2019-yj suggested that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered. @chicco2010renyi addressed information theory-based clustering such as Shannon or Renyi entropy and its variations. @Melnykov2013-sp discussed how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions. None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.

@@

```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for our method for clustering time series.", out.width="100%"}
knitr::include_graphics("img/flowchart-comment.png")

@@

(ref:hod-ind-group-png) The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots. They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer. The number indicates a unique customer id and a prototype id. In each of the plots, the line
represents the median, and the grey region shows the area between the 25th and 75th percentiles.
